### 2024-09-21 project/progress_summary.tex
Added progress report and in-progress report.

\documentclass{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{cite}

\graphicspath{{project/paper_images/}}

\title{Deep Learning Project - Progress Summary}
\author{
    Group Members: \\
    Farooq Mahmud
}
\date{September 20, 2024}

\begin{document}

\maketitle

\section{Progress summary}
\begin{itemize}
    \item Read papers cited in working version of report.
    \item Downloaded subset of Imagenet images that mapped to CIFAR-10 image classes. 
    \item Scaled images to 224 x 244 which matches the input size of the Szegedy paper.
    \item Trained a simple Sequential model using downloaded images.
    \item Reading Tensorflow documentation on how to create the more complex model in the Szegedy paper.
\end{itemize}

\end{document}